---
title: Final Presentation
toc: true
toc-location: right
execute:
  echo: false
  warning: false
  message: false
params:
  number: 4
  weight: '6%'
  purpose: 'purpose'
---

```{r}
#| label: setup

source(here::here("_common.R"))

rubric <- read_csv(here::here('rubrics', '4-final-presentation.csv'))
maxPoints <- rubric %>%
    filter(rating == "Excellent") %>%
    summarise(max = sum(maxPoints)) %>%
    pull(max)
```

```{r child = here::here("fragments", "project.qmd")}
```

Make a slideshow presentation of your findings in your final report. You may do one of the following: 

- Create a video recording of your presentation and publish it on the web (e.g. via Youtube, Vimeo, etc.).
- Present your presentation live in class.

We will watch all the presentations together during the period scheduled for the exam (December 13, 2023).

Below is a list of specific items your presentation should include.

## 1. Format

- Your presentation should be no longer than 10 minutes in length (it can be shorter though).
- Each team member must present at least one slide.
- You may use as many slides as you feel helps you communicate your ideas, but keep the 10 minute limit in mind (1-2 slides / min is a good guideline).
- Your title slide should include the project title, team member names, and the presentation date.
- All slides should be numbered in the bottom-left or bottom-right corners.

## 2. Recording strategies

If you choose to record your presentation, I recommend using Zoom. Have someone share their screen showing the slides and talk over the slides while recording the call.

Another option is to use a screen recording program, such as Quicktime (mac only), to record a screen recording on someone talking over the slides, though this may be challenging if not all teammates can be in the same room.

## 3. Content

Your slides do not need to include the detailed code used to conduct your analysis - that should be accessible from your report. Rather, the purpose of your presentation is to present the "big picture" overview of your project. You should discuss:

1. What you studied and why it matters (i.e. your research question).
2. What data you used (describe your sources)
3. What you found (show polished charts, discuss key results).
4. Any conclusions or new hypotheses you learned from your analyses.

## 4. Publish 

If you recorded a video, publish your video recording on the web (e.g. via Youtube, Vimeo, etc.), then **send a link to the published video in a DM to Professor Helveston on slack.**

Only one team member needs to publish the video.

## 5. Submit

Create a zip file of your slides as a pdf / pptx / qmd (whatever you used to make your slides) and submit it on the corresponding assignment submission on Blackboard. **Only one person from your team should submit.**

## Grading Rubric

### `r maxPoints` Total Points {.unnumbered}

```{r}
make_rubric(rubric)
```
