---
title: "Lab 5"
author: "Your Name"
date: "2025-XX-XX"
params:
  number: 5
  purpose: "The purpose of this lab is to follow a public data tutorial and recreate a water insecurity visualization using Census and spatial data."
format: 
  html:
    toc: true
    toc-location: right
    code-fold: true
    theme: flatly
    self-contained: true
execute:
  echo: true
  warning: false
  message: false
---


```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    fig.path = "figs/",
    fig.width = 7.252,
    fig.height = 4,
    comment = "#>",
    fig.retina = 3
)
```

## Before your start the lab notebook

- Update the header - put your name in the `author` argument and put today's date in the `date` argument.
- Click the "Render" button in RStudio and then open the rendered `5-lab5.html` page.
- Then go back and try changing the `theme` argument in the header to something else - you can see other available themes [here](https://quarto.org/docs/output-formats/html-themes.html). Notice the difference when you render now!

## Lab 5: Mapping Water Insecurity with `tidycensus` and `ggplot2`


In this lab, you will recreate a public map visualization of plumbing access across the U.S. using data from the American Community Survey (ACS) via the `tidycensus` package. You will:

- Use `get_acs()` to retrieve county-level data
- Clean and reshape the data
- Visualize the spatial distribution of plumbing access using `ggplot2`
- Create a choropleth map
- Complete a short design challenge to create your own variation of the map

### Reference Tutorial:
[Mapping water insecurities in R with tidycensus](https://waterdata.usgs.gov/blog/acs-maps/#visualizing-total-population-vs-percent-lacking-plumbing-across-counties-in-the-western-us)

### Suggested Reading:
- [R for Data Science (2e), Chapter 11 – Visualize](https://r4ds.hadley.nz/visualize)

---


## 1 Load Libraries & Set Up

```{r}
library(tidyverse)
library(tidycensus)
library(sf)
library(ggthemes)
library(viridis)

# In .Renviron
# CENSUS_API_KEY="your census"

census_api_key(Sys.getenv("CENSUS_API_KEY"), install = TRUE, overwrite = TRUE)

```


## 2 Acquire the Data

Use get_acs() to retrieve county-level data for the Western states in 2022 and 2023:

- Total population: `B01003_001`

- Households without plumbing: `B25016_003`

```{r}
# your code here, or use the code from tutorial
```

## 3 Pivot the data wider and calculate the percent of people without plumbing

Pivot the data wider so that each row contains both population and plumbing information. Check your result using `glimpse()` or `summary()`.

```{r}
# your code here, or use the code from tutorial
```


## 4 Create a choropleth map showing % without plumbing.

In this tutorial, they use `geom_sf()` to make choropleth map, but you can also use `tmap()` like we covered in the lecture if you prefer.

```{r}
# your code here, or use the code from tutorial
```

What does this preview map show? Are there clear spatial patterns?

::: {.callout-note title="Your Answer"}
Type your response here.
:::

## 5 Refined Choropleth Map

Recreate the polished map from the USGS tutorial. Consider:

- Restricting to western U.S. states (AK, AZ, CA, CO, HI, ID, MT, NM, NV, OR, UT, WA, WY)
- Adjusting the color scale
- Adding meaningful titles and captions

```{r}
# your code here, or use the code from tutorial
```

What refinements did you add in step 5 compared to step 4? How do these changes improve the visualization?

::: {.callout-note title="Your Answer"}
Type your response here.
:::

## 6 Create Your Own Map

Use the same dataset and design a different map. Possibilities include:

- Plotting absolute number of people without plumbing
- Faceting by region or state or urban/rural classification
- Changing map projection or styling

```{r}
# your code here
```

::: {.callout-note title="Your Interpretation"}
explain what your new visualization shows.
:::


## Save and Push Your Work

Remember to save your `.qmd` and render the HTML output before committing to GitHub.

```{bash eval = FALSE}
git add 5-lab5.qmd 5-lab5.html
git commit -m "Final submission for Lab5"
git push
```

