---
title: "Trends"
subtitle: "EMSE 4572 / 6572: Exploratory Data Analysis"
format: html
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    fig.path = "figs/",
    fig.width = 7.252,
    fig.height = 4,
    comment = "#>",
    fig.retina = 3
)

# # Install new packages (only do this once!)
# install.packages('viridis')
# install.packages('gganimate')
# install.packages('magick')

library(cowplot)
library(tidyverse)
library(viridis)
library(ggrepel)
library(gganimate)
library(magick)
library(here)


# Load all the data sets:
global_temps <- read_csv(here::here('data', 'nasa_global_temps.csv'))
us_covid     <- read_csv(here::here('data', 'us_covid.csv'))
```

# 1. Visualizing single variables over time

Use the `global_temps` data frame to explore ways to visualize the change in average global temperatures.

Consider using:

- points
- lines
- areas
- smoothed lines

## Points

```{r}
#| label: global-temps-points
#| fig.height: 5
#| fig.width: 6

# Write code here
```

## Line

```{r}
#| label: global-temps-line
#| fig.height: 5
#| fig.width: 6

# Write code here
```

## Smoothed line

```{r}
#| label: global-temps-line-smoothed
#| fig.height: 5
#| fig.width: 6

# Write code here
```

## Area

```{r}
#| label: global-temps-area
#| fig.height: 5
#| fig.width: 6

# Write code here
```



# 2. Animation

Use the `global_temps` data frame to animate one of the charts we did in the first exercise on changes in average global temperatures.

## Points

```{r}
#| label: global-temps-points-anim
#| fig.height: 5
#| fig.width: 6

# Write code here
```

## Line

```{r}
#| label: global-temps-line-anim
#| fig.height: 5
#| fig.width: 6

# Write code here
```

## Area

```{r}
#| label: global-temps-area-anim
#| fig.height: 5
#| fig.width: 6

# Write code here
```



# 3. Visualizing multiple variables over time

Use the `us_covid` data frame to explore ways to visualize the number of daily cases using:

1. A labeled line chart
2. A stacked area chart
3. A heat map

## Labeled line chart

For this one, try showing just the lines for NY and CA.
I already made the summary data frame to do this.

```{r}
#| label: covid-ca-ny
#| echo: false
#| fig.height: 4
#| fig.width: 12
#| fig.align: 'center'

us_covid_ca_ny <- us_covid %>%
  mutate(cases_daily = cases_daily / 10^3) %>%
  filter(state %in% c("California", "New York"))

# Write code here
```

## Stacked area chart

For this one, try showing NY, CA, and all other states.
I already made the summary data frame to do this.

```{r}
#| label: covid-area
#| fig.height: 4
#| fig.width: 12
#| fig.align: 'center'

us_covid %>%
  mutate(
    state = fct_other(state, keep = c(
      "California", "New York")),
    state = fct_relevel(state, c(
      "Other", "California", "New York"
    ))) %>%
  group_by(day, state) %>%
  summarise(deaths_daily = sum(deaths_daily))

# Write code here
```

```{r}
#| label: covid-area-labeled
#| fig.height: 4
#| fig.width: 12
#| fig.align: 'center'

# Write code here
```

## Heat map

```{r}
#| label: covid-heatmap
#| fig.height: 7
#| fig.width: 10
#| fig.align: 'center'

# Write code here
```
